---
title: "eitan-q4"
output: html_document
---

## Part 4 — Simpler Panel (will use LASSO)

To generate a simpler panel, we will use LASSO in order to zero out unimportant coefficients, which hopefully will result in a smaller set of proteins while retaining accuracy. We first will tune it on the training set only, then evaluate once on the test set.

First, we run setup the libraries, load the data, and set a seed for reproducability. 

``` {r}

## ---- part4-lasso-simple, message=FALSE, warning=FALSE ------------------
library(glmnet)
library(pROC)
library(caret)

# Load data (this file is one folder up, inside 'data/')
load("../../data/biomarker-clean.RData")

set.seed(133233)
```


Now, we split the data, so we do not end up with training on all the data and overfitting on that data.
```{r}
idx  <- caret::createDataPartition(biomarker_clean$group, p = 0.8, list = FALSE)
train <- biomarker_clean[idx, ]
test  <- biomarker_clean[-idx, ]

train$group <- factor(train$group, levels = c("TD","ASD"))
test$group  <- factor(test$group,  levels = c("TD","ASD"))

x_train <- model.matrix(group ~ . - 1, data = subset(train, select = -ados))
y_train <- train$group
x_test  <- model.matrix(group ~ . - 1, data = subset(test,  select = -ados))
y_test  <- test$group
```

Now, we fit a LASSO logistic regression and extract the panel by looking at the coefficients generated. 

```{r}
set.seed(133233)

cvfit <- cv.glmnet(
  x = x_train, y = y_train,
  family = "binomial",
  type.measure = "auc",
  nfolds = 8
)

coef_1se <- coef(cvfit, s = cvfit$lambda.1se)
selected <- rownames(coef_1se)[as.numeric(coef_1se) != 0]
selected <- setdiff(selected, "(Intercept)")

cat("Simpler panel size:", length(selected), "\n")
print(selected)
```

Now, we use the test data determined earlier we can test our selected panel to see how it compares. 
``` {r}
pred <- as.numeric(predict(cvfit, newx = x_test, s = cvfit$lambda.1se, type = "response"))
roc_obj <- roc(response = y_test, predictor = pred, levels = c("TD","ASD"), quiet = TRUE)
auc_val <- as.numeric(auc(roc_obj))
cat(sprintf("Test AUROC (simpler λ_1SE model): %.3f\n", auc_val))
```

Analysis why 0.636 is lower and explain todo